---
title: "Solids tumor cohort - complex svs analysis"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document: default
date: '2023-06'
editor_options: 
  chunk_output_type: console
---
<style type="text/css">
h1 { font-size: 24px; }
h2 { font-size: 18px; }
h3 { font-size: 14px; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      results = "hold",
                      warning = F,
                      message = F,
                      cache=T,
                      opts.label="kill_prefix",
                      fig.width=8, fig.height=5,
                      dev="png") 

```


```{r include=F}

suppressPackageStartupMessages({
  library(tidyverse, quietly=TRUE)
  library(stringr, quietly=TRUE)
  library(stringi)
  library(dplyr, quietly=TRUE)
  library(GenomicRanges)
  library(openssl)
  library(igraph)
  library(pheatmap)
  library(RColorBrewer)
  library(ggrepel)
  library(ggpattern)
})



root_dir="~/PycharmProjects/"
source(paste0(root_dir,"structuralvariation/sv_functional_analysis/default.conf"))

source(paste0(script_dir,"functions.general.R"))
source(paste0(script_dir,"functions.svs.R"))
source(paste0(script_dir,"functions.expression.R"))
source(paste0(script_dir,"functions.cna.R"))

source(paste0(wdir,"solids.conf"))


## Fill path templates
autosomes= c(paste("chr",1:22,sep=""),"chrX")
annotation_colors_call = c(`gain` = "red3", `loh` = "tan2", `loss` = "steelblue3", `amp`="red4")


plot_dir=cohort_results_dir

map_template_vars=c('${resources_dir}'=resources_dir,
                    '${merged_svs_dir}'=merged_svs_dir,
                    #'${cohort_wdir}'=cohort_wdir,
                    #'${cohort_identifier}'=cohort_identifier,
                    '${cna_data_dir}'=cna_data_dir,
                    '${cna_seg_file_ext}'=cna_seg_file_ext,
                    '${sequencing_strategy}'="WGS")

run_date=20230701
merged_svs_classes_path = paste0(cohort_results_dir,"merged_svs_classes.",run_date,".tsv")
merged_svs_classes_annotated_path = paste0(cohort_results_dir,"merged_svs_classes_annotated.",run_date,".tsv")
complex_svs_annotated_path = paste0(cohort_results_dir,"complex_svs_annotated.",run_date,".tsv")
complex_sv_classification_path = paste0(cohort_results_dir,"complex_sv_classification.",run_date,".tsv")
complex_svs_effect_path = paste0(cohort_results_dir,"complex_svs_annotated.",run_date,".effect.tsv")
driver_gene_alterations_path = paste0(cohort_results_dir,"clinrel_mutations_incl_prognostic.",run_date,".tsv")

tmb_path=paste0(cohort_results_dir,"cohort_snv_burden_somatic.",run_date,".tsv")
fga_path=paste0(cohort_results_dir,"cohort_fga_per_chrom.",run_date,".tsv")
complex_fga_path = paste0(cohort_results_dir,"cohort_fga_per_chrom.complex.",run_date,".tsv")
ploidy_path=paste0(cohort_results_dir,"cohort_ploidy.",run_date,".tsv")
cohort_fga_per_chrom_per_call_path = paste0(cohort_results_dir,"cohort_fga_per_chrom_per_call.",run_date,".tsv")

mutation_burden_path = paste0(cohort_results_dir,"mutation_burden.",run_date,".tsv")
clinical_events_enrichment_path = paste0(cohort_results_dir,"enrichment_results.clinical_events.",run_date,".tsv")
mutation_burden_enrichment_path = paste0(cohort_results_dir,"enrichment_results.mutation_burden.",run_date,".tsv")



```


```{r}

annotation_colors$complex_sv_classes = c(
`amplicon` = "#66c2a5",
`chromoplexy`="#fc8d62",
`chromothripsis`="#8da0cb",
`complex_other`="#e78ac3",
`sv_pair_ctx`="#a6d854",
`sv_pair`="#ffd92f",
`reciprocal_ctx`="#e5c494",
`simple`="#dddddd")


annotation_colors$complex_sv_superclasses = c(
`amplicon` = "#66c2a5",
`chromoplexy`="#fc8d62",
`chromothripsis`="#8da0cb",
`complex_other`="#e78ac3",
`simple`="#dddddd")

cancer_type.labs = c("EWS","NBL","fpRMS","fnRMS","WT","HBL")
names(cancer_type.labs) = c("ewing_sarcoma","neuroblastoma","alveolar_rhabdomyosarcoma","embryonal_rhabdomyosarcoma","nephroblastoma","hepatoblastoma")

```


# Functions

# Cohort 

Load cohort and annotate 

```{r} 
cohort = read.table(patient_table_path,sep = "\t", header=T) 
patient_tumor_id_cols = c("patient_label","tumor_id","cancer_type")

```


# Read data
## Load SVs > merged svs classes

```{r}
merged_svs = read.table(merged_svs_classes_annotated_path,sep="\t",header=T)
complex_svs_annotated = read.table(complex_svs_effect_path,sep="\t",header=T)
complex_sv_classification = read.table(complex_sv_classification_path,sep="\t",header=T)

```


## Driver alterations 

2023-09-03 moved making table clinrel_mutations_incl_prognostic to dedicated script (utils/make_driver_gene_alteration_overview.R)
```{r}

clinrel_mutations_incl_prognostic = read.table(driver_gene_alterations_path, sep="\t",header=T)
driver_gene_alteration = clinrel_mutations_incl_prognostic

driver_gene_alteration_complex_summary= clinrel_mutations_incl_prognostic %>% filter(complex_sv_alteration) %>% 
  group_by(patient_label,complex_sv_id) %>% summarize(driver_gene_lst=lst_str(gene_name),sv_driver_expected=any(sv_expected),complex_sv_driver_gene_expected=any(complex_sv_expected))

tp53_or_mdm2_mutations = clinrel_mutations_incl_prognostic  %>% filter(gene_name %in% c("TP53","MDM2"))

```

#  Make overview patient_level_complex_sv_overview
```{r}

complex_svs_per_patient = merged_svs %>% 
  filter(flag_is_complex) %>% 
  group_by(patient_label) %>% summarize(complex_svs_cnt=cnt_str(complex_sv_id),complex_classes_cnt=cnt_str(complex_sv_class_super),complex_classes_lst=lst_str(complex_sv_class_super)) %>% arrange(-complex_classes_cnt) %>% as.data.frame()

patient_level_complex_sv_overview = cohort  %>% 
  left_join(complex_svs_per_patient) %>% 
  dplyr::mutate(complex_sv = patient_label %in% filter(merged_svs,flag_is_complex)$patient_label,
                complex_svs_cnt= ifelse(is.na(complex_svs_cnt),0,complex_svs_cnt),
                complex_sv_driver_gene = patient_label %in% driver_gene_alteration_complex_summary$patient_label,
                sv_driver_expected = patient_label %in%
                  filter(driver_gene_alteration_complex_summary,sv_driver_expected)$patient_label,
                complex_sv_driver_gene_expected = patient_label %in%
                  filter(driver_gene_alteration_complex_summary,complex_sv_driver_gene_expected)$patient_label,
                driver_gene_mutated = patient_label %in% clinrel_mutations_incl_prognostic$patient_label,
                driver_gene_mutated_no_complex = patient_label %in% filter(clinrel_mutations_incl_prognostic,!complex_sv_alteration)$patient_label,
                 complex_sv_any_effect = patient_label %in% filter(complex_svs_annotated,effect!="unknown")$patient_label,
                tp53_or_mdm2 = patient_label %in% tp53_or_mdm2_mutations$patient_label)

complex_svs_driver_cnts = complex_svs_annotated %>% filter(!is.na(driver_gene_lst)) %>% group_by(patient_label) %>% summarize(complex_svs_cnt_driver=cnt_str(complex_sv_id))

if("complex_svs_cnt_driver" %in% names(patient_level_complex_sv_overview)) {
patient_level_complex_sv_overview = patient_level_complex_sv_overview %>% select(-complex_svs_cnt_driver) 
}
patient_level_complex_sv_overview = patient_level_complex_sv_overview %>% left_join(complex_svs_driver_cnts)
patient_level_complex_sv_overview[is.na(patient_level_complex_sv_overview$complex_svs_cnt_driver),c("complex_svs_cnt_driver")]=0

patient_level_complex_sv_overview$remark=""
tumors_with_connected_amplicons = c("M962AAB","M959AAB","M157AAB")
patient_level_complex_sv_overview[patient_level_complex_sv_overview$patient_label %in% tumors_with_connected_amplicons,c("remark")]="Two complex SV clusters connected by low AF SVs so single event"


```

## export supplementary table

```{r}
cohort  = cohort %>% left_join(patient_level_complex_sv_overview %>% select(patient_label, driver_gene_mutated, complex_sv, complex_sv_driver_gene))

#should be sufficient to perform analysis => the complex sv annotation added in analysis document

write.table(cohort,paste0(metadata_dir,"cohort.solids_v4.20230630.supplementary_table.tsv"),sep="\t",col.names = T,row.names = F,quote=T)

```
# Mutation burden

## quantification svs complex svs
```{r}
quantification_svs = merged_svs
quant_svs_total = quantification_svs %>% cnt_uq(uq_id="patient_sv_merged",group_cols = c("patient_label"),attr_name = "total")

label_order_svs = quant_svs_total %>% arrange(-total) %>% select(patient_label) %>% flatten_chr()
label_order_svs = c(label_order_svs, cohort %>% filter(!patient_label %in% label_order_svs) %>% select(patient_label) %>% flatten_chr())


```

```{r}
##super classes count svs
## different classes of complex 
superclasses = quantification_svs$complex_sv_class_super %>% unique()
superclasses = superclasses[!is.na(superclasses)]

quant_svs_superclasses = cohort[,c("patient_label"),drop=F]

for(complex_class in superclasses) {
  quant_svs_superclasses = quant_svs_superclasses %>% left_join( 
  quantification_svs %>% filter(complex_sv_class_super==complex_class) %>% cnt_uq(uq_id="patient_sv_merged",group_cols = c("patient_label"),attr_name = complex_class) ) 
}
  
quant_svs_superclasses_wide = quant_svs_superclasses

quant_svs_superclasses = quant_svs_superclasses_wide %>% pivot_longer(-patient_label,values_to = "cnt",names_to = "attr") %>% left_join(cohort[,c("patient_label","cancer_type")])


quant_svs_superclasses$patient_label=factor(quant_svs_superclasses$patient_label,levels=label_order_svs)
quant_svs_superclasses[is.na(quant_svs_superclasses)]=0


```

```{r}

## counting complex events
quant_complex_superclasses = cohort[,c("patient_label"),drop=F]  

for(complex_class in superclasses) {
  quant_complex_superclasses = quant_complex_superclasses %>% left_join( 
  quantification_svs %>% filter(complex_sv_class_super==complex_class) %>% cnt_uq(uq_id="complex_sv_id",group_cols = c("patient_label"),attr_name = complex_class) ) 
}

quant_complex_superclasses_wide = quant_complex_superclasses


```

## make mutation burden table


```{r}
cohort_tmb = read.table(tmb_path,sep="\t",header=T)
cohort_tmb = cohort_tmb %>% left_join(cohort[,c("patient_label","cancer_type","sample_days_after_diagnosis")])

cohort_fga = read.table(fga_path,sep="\t",header = T) 
cohort_fga = cohort_fga %>% filter(seqnames=="total")
cohort_fga = cohort_fga %>% left_join(cohort[,c("patient_label","cancer_type")])

complex_fga = read.table(complex_fga_path,sep="\t",header = T) 
complex_fga = complex_fga %>% filter(seqnames=="total")
complex_fga = complex_fga %>% left_join(cohort[,c("patient_label","cancer_type")])

cohort_ploidy = read.table(ploidy_path,sep="\t",header=T)

cohort_fga_per_chrom_per_call = read.table(cohort_fga_per_chrom_per_call_path,sep="\t",header = T) 
cohort_fga_per_chrom_per_call = cohort_fga_per_chrom_per_call %>% filter(seqnames=="total") %>% left_join(cohort[,c("patient_label","cancer_type")])

cohort_fga_gain = cohort_fga_per_chrom_per_call %>% filter(cn_type=="gain")
cohort_fga_loss = cohort_fga_per_chrom_per_call %>% filter(cn_type=="loss")

mutation_burden_overview_cols = c("patient_label","cancer_type","cnt_type","cnt","attr")
mutation_burden_overview = rbind_no_colmatch(
  quant_svs_superclasses %>% dplyr::mutate(cnt_type="1.complex_svs") %>% select(mutation_burden_overview_cols),
  cohort_tmb %>% dplyr::mutate(cnt_type="2.snv_indel",cnt=snv_indel_somatic_nonsyn_cnt,attr="snv_indel_somatic_nonsyn_cnt") %>% select(mutation_burden_overview_cols))

mutation_burden_overview = rbind_no_colmatch(mutation_burden_overview,
                                  cohort_fga_gain %>% dplyr::mutate(cnt_type="3.fga",cnt=fga,attr="fga_gain")  %>% select(mutation_burden_overview_cols))

mutation_burden_overview = rbind_no_colmatch(mutation_burden_overview,
                                  cohort_fga_loss %>% dplyr::mutate(cnt_type="3.fga",cnt=fga,attr="fga_loss")  %>% select(mutation_burden_overview_cols))

mutation_burden_overview = rbind_no_colmatch(mutation_burden_overview,
                                  cohort_ploidy %>% dplyr::mutate(cnt_type="4.ploidy",cnt=custom_ploidy,attr="ploidy")  %>% select(mutation_burden_overview_cols))

mutation_burden_overview = rbind_no_colmatch(mutation_burden_overview,
                                  complex_fga %>% dplyr::mutate(cnt_type="5.complex_fga",cnt=fga_complex,attr="complex_fga")  %>% select(mutation_burden_overview_cols))

mutation_burden_overview = rbind_no_colmatch(mutation_burden_overview,
cohort_tmb %>% dplyr::mutate(cnt_type="6.tmb",cnt=snv_indel_somatic_nonsyn_cnt,attr="tmb") %>% select(mutation_burden_overview_cols))


mutation_burden_overview = rbind(mutation_burden_overview, tp53_or_mdm2_mutations %>% dplyr::mutate(attr=gene_name,cnt=NA,cnt_type="1.complex_svs") %>% select(mutation_burden_overview_cols))

mutation_burden_overview$patient_label=factor(mutation_burden_overview$patient_label,levels=label_order_svs)
#mutation_burden_overview[is.na(mutation_burden_overview)]=0

```

### export

```{r}
mutation_burden_overview_wide = mutation_burden_overview %>% pivot_wider(id_cols = c("patient_label","cancer_type"), names_from="attr",values_from = "cnt")
mutation_burden_overview_wide = mutation_burden_overview_wide %>% left_join(quant_svs_total %>% dplyr::rename(svs_cnt=total)) %>% left_join(patient_level_complex_sv_overview %>% select(patient_label,complex_svs_cnt,clinical_event) %>% dplyr::rename(complex_events_cnt=complex_svs_cnt) )
#events vs svs 
mutation_burden_overview_wide = mutation_burden_overview_wide %>% dplyr::rename(ecdna_amplicon=amplicon) 

mutation_burden_overview_wide_cols = c("patient_label","cancer_type","svs_cnt","complex_events_cnt")
mutation_burden_overview_wide_cols = c(mutation_burden_overview_wide_cols, names(mutation_burden_overview_wide)[!names(mutation_burden_overview_wide) %in% mutation_burden_overview_wide_cols])

mutation_burden_overview_wide = mutation_burden_overview_wide[,mutation_burden_overview_wide_cols]
write.table(mutation_burden_overview_wide,mutation_burden_path,sep="\t",quote=F,row.names = F,col.names = T)
```

## simple figure

```{r}
if(FALSE){
pdf(paste0(cohort_results_dir,"complex_svs_quantification.pdf"),width=18,height=5)

p = ggplot(quant_svs_superclasses,aes(x=patient_label,y=cnt,fill=attr)) + geom_col() +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90)) +
  scale_fill_manual(values=annotation_colors$complex_sv_superclasses ) +
  ggtitle("#SVs: complex classes, superclasses")

p +   facet_grid(cols=vars(cancer_type),scales="free",space = "free")


## counting complex 

quant_complex_superclasses = quant_complex_superclasses_wide %>% pivot_longer(-patient_label,values_to = "cnt",names_to = "attr") %>% left_join(cohort[,c("patient_label","cancer_type")])

quant_complex_superclasses$patient_label=factor(quant_complex_superclasses$patient_label,levels=label_order_svs)

#combine counting svs and counting complex events to show how they collapse

quant_complex_superclasses = rbind(quant_svs_superclasses %>% dplyr::mutate(cnt_type="2.complex_svs"),
                                  quant_complex_superclasses %>% dplyr::mutate(cnt_type="1.complex_events"))

quant_complex_superclasses$patient_label=factor(quant_complex_superclasses$patient_label,levels=label_order_svs)


p = ggplot(quant_complex_superclasses %>% filter(attr != "simple"),aes(x=patient_label,y=cnt,fill=attr)) + geom_col() +
  geom_hline(yintercept = 0,color="black") +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90)) +
  scale_fill_manual(values=annotation_colors$complex_sv_superclasses ) +

  ggtitle("Comparing #svs and #complex events, per superclass")

p_freex = p  +   facet_grid(cols=vars(cancer_type),rows=vars(cnt_type),scales="free",space = "free_x") 
p_freex

dev.off()

}
```

## manuscript figure: mutation burden with TMB and FGA as bars 
2023-07-14 

```{r}
mutation_burden_overview$cancer_type = factor(mutation_burden_overview$cancer_type,levels = names(cancer_type.labs))
```

```{r}
pdf(paste0(cohort_results_dir,"complex_svs_quantification.tmb_fga.pdf"),width=18,height=9)
#pdf(paste0(cohort_results_dir,"complex_svs_quantification.tmb_fga.squished.pdf"),width=18,height=4)

#pdf(paste0(cohort_results_dir,"complex_svs_quantification.tmb_fga_ploidy_complex.pdf"),width=18,height=9)

p = ggplot(mutation_burden_overview %>% filter(cnt_type!="4.ploidy" & cnt_type != "5.complex_fga" & cnt_type != "6.tmb" & cnt>0) ,aes(x=patient_label,y=cnt,fill=attr)) + 
  geom_col(color="white",size=0) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values=c(annotation_colors$complex_sv_superclasses,`fga_gain`="red3",`fga_loss`="steelblue3") ) +
  ggtitle("#SVs: complex classes") +
  facet_grid(cols=vars(cancer_type),rows=vars(cnt_type),scales="free",space = "free_x") 

p = p + geom_point(data=filter(mutation_burden_overview,is.na(cnt)),y=-6,aes(shape=attr),size=2)

p +   facet_grid(cols=vars(cancer_type),rows=vars(cnt_type),scales="free",space = "free_x",labeller = labeller(cancer_type = cancer_type.labs)) + theme(legend.position="bottom")#, legend.direction = "vertical")

p



p = ggplot(mutation_burden_overview ,aes(x=patient_label,y=cnt,fill=attr)) + 
    geom_col(color="white",size=0) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values=c(annotation_colors$complex_sv_superclasses,`fga_gain`="red3",`fga_loss`="steelblue3") ) +
  ggtitle("#SVs: complex classes") +
  facet_grid(cols=vars(cancer_type),rows=vars(cnt_type),scales="free",space = "free_x") 

p + geom_point(data=filter(mutation_burden_overview,is.na(cnt)),y=-6,aes(shape=attr),size=2)


dev.off()
```

```{r}
pdf(paste0(cohort_results_dir,"fga_complex.pdf"),width=18,height=3)

plot_complex_fga = cohort[,c("patient_label","cancer_type")] %>% left_join(complex_fga) #get all patients
complex_fga$fga_complex

plot_complex_fga$patient_label=factor(plot_complex_fga$patient_label,levels=label_order_svs)

p = ggplot(plot_complex_fga ,aes(x=patient_label,y=fga_complex,fill=fga_complex_relative)) + 
  geom_col(color="white",size=0) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90)) +
  scale_fill_continuous(type="viridis") +
  facet_grid(cols=vars(cancer_type),scales="free",space = "free_x") 
p

dev.off()
```

## enrichment

```{r}
mutation_burden_annotated=mutation_burden_overview_wide %>% left_join(patient_level_complex_sv_overview %>% select(patient_label,complex_sv,tp53_or_mdm2,days_sample_before_treatment))
mutation_burden_annotated$fga=mutation_burden_annotated$fga_gain+mutation_burden_annotated$fga_loss

mutation_burden_annotated = mutation_burden_annotated %>% mutate(treated_sample = days_sample_before_treatment<(-14))

mutation_burden_metrics = c("fga","fga_gain","tmb","snv_indel_somatic_nonsyn_cnt","complex_events_cnt","svs_cnt")
mutation_categories = c("complex_sv","tp53_or_mdm2","treated_sample")
mutation_burden_enrichment=data.frame()

for(ct in mutation_categories) {
for(mb in mutation_burden_metrics){
  row=c()
  ct_values = unique(mutation_burden_annotated[,ct]) %>% length()
  if(ct_values>2) {
  test = kruskal.test(data=mutation_burden_annotated, formula(paste0(mb," ~ ",ct)))
  row$type="kruskal"
  } else {
  test = wilcox.test(data=mutation_burden_annotated, formula(paste0(mb," ~ ",ct)))
  row$type="wilcox"
  }
  row$pval = test$p.value
  row$ct=ct
  row$mb=mb
  medians = mutation_burden_annotated %>% group_by(across(all_of(ct))) %>% get_min_max_median(mb) %>% select(c(ct,"median")) %>% pivot_wider(names_from=ct,values_from="median")
  row$median_true=medians$`TRUE`
  row$median_false=medians$`FALSE`
  median_overall = mutation_burden_annotated %>% get_min_max_median(mb) 
  row$median = median_overall$median
  mutation_burden_enrichment = rbind(mutation_burden_enrichment,row)
}
}

mutation_burden_enrichment=mutation_burden_enrichment %>% arrange(ct,pval)%>% mutate(sign=pval<0.05)
mutation_burden_enrichment %>% filter(ct=="tp53_or_mdm2")
mutation_burden_enrichment %>% filter(ct=="complex_sv")
mutation_burden_enrichment %>% filter(ct=="treated_sample")


write.table(mutation_burden_enrichment,enrichment_results_mutation_burden_path,sep="\t",row.names = F,col.names = T,quote=F)

```

```{r}

mutation_burden_annotated %>% get_min_max_median("svs_cnt")
mutation_burden_annotated %>% filter(complex_sv) %>% get_min_max_median("complex_events_cnt")
mutation_burden_annotated %>% filter(complex_sv&tp53_or_mdm2) %>% get_min_max_median("complex_events_cnt")




```


#  Clinical events / relapse association

```{r}

complex_sv_clinical = patient_level_complex_sv_overview 
counts_col_lst = c("clinical_events_cnt")
for(col in counts_col_lst) {
  complex_sv_clinical[is.na(complex_sv_clinical[,col]),col]=0
}

clinical_events_enrichment=data.frame()


test_under_enriched=F
filter_column_lst=c("complex_sv","complex_sv_driver_gene","sv_driver","tp53_or_mdm2","driver_gene_mutated","complex_sv_any_effect","driver_gene_mutated_no_complex")
split_attr_lst=c("cancer_type","driver_gene_mutated")

names(complex_sv_clinical)

filter_column_lst = filter_column_lst[filter_column_lst %in% names(complex_sv_clinical)]
counts_col_lst = counts_col_lst[counts_col_lst %in% names(complex_sv_clinical)]
split_attr_lst = split_attr_lst[split_attr_lst %in% names(complex_sv_clinical)]


for(filter_column in filter_column_lst) {
for(counts_col in counts_col_lst) {
for(counts_col_threshold in c(1)){

for(split_attr in split_attr_lst) {
split_value_lst = unique(complex_sv_clinical[,split_attr])
if(split_attr=="cancer_type") { split_value_lst=c("pan_cancer",split_value_lst)}
for(split_value in split_value_lst) {
  if(split_value=="pan_cancer") {
    event_counts_selection = complex_sv_clinical
  } else {
    event_counts_selection = complex_sv_clinical %>% filter(!!sym(split_attr) == split_value)
  }
  
  under_and_condition = uq_patients(event_counts_selection %>% filter(!!sym(counts_col)<counts_col_threshold) %>% filter(if_all(filter_column)))
  under_not_condition = uq_patients(event_counts_selection %>% filter(!!sym(counts_col)<counts_col_threshold)%>% filter(!if_any(filter_column)))
  over_and_condition = uq_patients(event_counts_selection %>% filter(!!sym(counts_col)>=counts_col_threshold) %>% filter(if_all(filter_column)))
  over_not_condition = uq_patients(event_counts_selection %>% filter(!!sym(counts_col)>=counts_col_threshold) %>% filter(!if_any(filter_column)))
              
  if(test_under_enriched) {
  fisher.table= matrix(c(  under_and_condition,  under_not_condition,  
                           over_and_condition,  over_not_condition   ),nrow=2)
  } else {
    fisher.table= matrix(c(  over_and_condition,  over_not_condition,  
                             under_and_condition,  under_not_condition ),nrow=2)
  }
  
  print(paste0(split_value, " ",counts_col," ",counts_col_threshold))
  print(fisher.table)
  t=fisher.test(fisher.table)
  print(t)
  
  #case is under thresh meets condition and control over threshold meets condition
  
  row=c()
  row$test_under_enriched=test_under_enriched
  row$filter_column=filter_column
  row$split_attr=split_attr
  row$split_value=split_value
  row$counts_col = counts_col
  row$counts_col_thresh = counts_col_threshold
  row$pval=t$p.value
  row$odds_ratio=t$estimate[[1]]
  row$conf_int=toString(c(t$conf.int[1],t$conf.int[2]))
  
  row$frac_under_and_condition_of_under=under_and_condition/(under_and_condition+under_not_condition)
  row$frac_over_and_condition_of_over=over_and_condition/(over_and_condition+over_not_condition)
  row$frac_under_and_condition_of_condition=under_and_condition/(under_and_condition+over_and_condition)
  row$frac_over_and_condition_of_condition=over_and_condition/(under_and_condition+over_and_condition)
  row$frac_under_and_not_condition_of_not_condition=under_not_condition/(under_not_condition+over_not_condition)
  row$frac_over_and_not_condition_of_not_condition=over_not_condition/(under_not_condition+over_not_condition)

  row$under_and_condition=under_and_condition
  row$over_and_condition=over_and_condition
  row$under_and_not=under_not_condition
  row$over_and_not=over_not_condition
  

  clinical_events_enrichment=rbind(clinical_events_enrichment,row)
} 
}
}
}
}


write.table(clinical_events_enrichment,clinical_events_enrichment_path,sep = "\t",col.names = T,row.names = F,quote=F)

```
## analyse enrichment results
```{r}
#original:
clinical_events_enrichment %>% filter(split_value=="pan_cancer") %>% select(-contains("frac_under"),-conf_int,-counts_col_thresh,-test_under_enriched) %>% arrange(pval) %>%
  filter(counts_col=="clinical_events_cnt")
  #filter(filter_column=="complex_sv")
  #filter(filter_column=="complex_sv_any_effect" )
  #filter(filter_column=="complex_sv_driver_gene" )

```


Sanity checks:
```{r}
complex_sv_clinical %>% nrow()
complex_sv_clinical %>% filter(clinical_event>=1) %>% nrow()
complex_sv_clinical %>% filter(complex_sv) %>% nrow()
complex_sv_clinical %>% filter(clinical_event>=1 & complex_sv) %>% nrow()
complex_sv_clinical %>% filter(complex_sv_any_effect) %>% nrow()
complex_sv_clinical %>% filter(complex_sv_driver_gene) %>% nrow()
complex_sv_clinical %>% filter(clinical_event>=1 & complex_sv_any_effect) %>% nrow()

complex_sv_clinical %>% filter(tp53_or_mdm2) 
```

2023-08-17 updated the gene list 

41% of tumors with a complex sv affecting driver gene had a clinical event compared to 19% of non-complex (p<0.05, OR2.8) 

complex sv detected remains the same
pan cancer => 2.45 enrichment pval = 0.056
63% of clinical event have complex sv, and 34% of complex sv had clinical event compared to 17% of non-complex. 
(19 pt both clinical and complex, of the 56 complex and total 30 clinical)



# Figure: complex events effect

```{r}
quant_complex_svs_effect = complex_svs_annotated %>% group_by(patient_label) %>% summarize(complex_svs_cnt_effect=sum(effect!="unknown"))

if("complex_svs_cnt_effect" %in% names(patient_level_complex_sv_overview)) {
  patient_level_complex_sv_overview = patient_level_complex_sv_overview %>% select(-complex_svs_cnt_effect)
}
patient_level_complex_sv_overview = patient_level_complex_sv_overview %>% left_join(quant_complex_svs_effect) %>% mutate(complex_svs_cnt_effect=ifelse(is.na(complex_svs_cnt_effect),0,complex_svs_cnt_effect))


label_order_complex_driver = patient_level_complex_sv_overview %>% arrange(complex_sv_driver_gene,driver_gene_mutated,tp53_or_mdm2,complex_svs_cnt_driver,complex_svs_cnt,clinical_event) %>% dplyr::mutate(patient_label = as.character(patient_label)) %>% select(patient_label) %>% unique() %>% flatten_chr() %>% rev()

#maybe sort of number of effect as well?
label_order_complex_effect = patient_level_complex_sv_overview %>% arrange(complex_sv_driver_gene,complex_svs_cnt_driver,complex_svs_cnt_effect,complex_svs_cnt,driver_gene_mutated,tp53_or_mdm2,clinical_event) %>% dplyr::mutate(patient_label = as.character(patient_label)) %>% select(patient_label) %>% unique() %>% flatten_chr() %>% rev()

write.table(label_order_complex_driver,paste0(cohort_results_dir,"label_order_complex_driver.tmp"))
write.table(label_order_complex_effect,paste0(cohort_results_dir,"label_order_complex_effect.tmp"))

patient_level_complex_sv_overview$patient_label=factor(patient_level_complex_sv_overview$patient_label,levels=label_order_complex_driver)
patient_level_complex_sv_overview$patient_label=factor(patient_level_complex_sv_overview$patient_label,levels=label_order_complex_effect)

complex_svs_annotated = complex_svs_annotated %>% left_join(patient_level_complex_sv_overview)
complex_svs_annotated$patient_label=factor(complex_svs_annotated$patient_label,levels=label_order_complex_driver)
complex_svs_annotated = complex_svs_annotated %>% arrange(effect)

```



with pattern
```{r}

display_patient_level_complex_sv_overview=patient_level_complex_sv_overview
display_patient_level_complex_sv_overview$cancer_type = factor(display_patient_level_complex_sv_overview$cancer_type,levels = names(cancer_type.labs))

display_complex_svs_annotated = complex_svs_annotated
display_complex_svs_annotated$cancer_type = factor(display_complex_svs_annotated$cancer_type,levels = names(cancer_type.labs))

if(FALSE){
p = ggplot(display_patient_level_complex_sv_overview) + 
  geom_point(aes(x=patient_label,y=1),color="white") +
  geom_point(data=filter(display_patient_level_complex_sv_overview,driver_gene_mutated),aes(x=patient_label,y=-0.25,color="clinrel")) +
  geom_point(data=filter(display_patient_level_complex_sv_overview,tp53_or_mdm2),aes(x=patient_label,y=-0.5,color="tp53/mdm2")) +
  geom_bar(data=display_complex_svs_annotated,aes(x=patient_label,fill=effect)) +
  #geom_col(color="white",size=0,aes(x=patient_label,y=cnt,fill=attr)) +
  facet_grid(cols=vars(cancer_type),scales="free",space = "free") + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values=c(`unknown`="grey",`driver`="red3",`target_gene_cn_change`="pink3",chrom_alt="steelblue3")) +
  ggtitle("#Complex events")

p = p + facet_grid(cols=vars(cancer_type),scales="free",space = "free",labeller = labeller(cancer_type = cancer_type.labs))
p 

p +   geom_point(data=filter(display_patient_level_complex_sv_overview,clinical_event),aes(x=patient_label,y=-0.75,color="clinical_event")) 
}

```
```{r}
pdf(paste0(cohort_results_dir,"complex_svs_effect.",run_date,".pdf"),width=18,height=5)

annotation_pattern_effect = c(`unknown`="none",`driver`="weave",`driver_and_chrom_alt`="circle",chrom_alt="stripe")
   
p = ggplot(display_patient_level_complex_sv_overview,aes(x=patient_label)) + 
  geom_point(aes(y=1),color="white") +
  geom_bar_pattern(data=display_complex_svs_annotated,aes(fill=complex_sv_class_super,pattern=effect,pattern_spacing=cancer_type,alpha=effect),pattern_color="black",color="white",pattern_fill="black",pattern_alpha=0.4) +
  facet_grid(cols=vars(cancer_type),scales="free_x",space = "free") + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values=annotation_colors$complex_sv_superclasses) +
  scale_pattern_manual(values=annotation_pattern_effect,guide="none") + 
  scale_pattern_spacing_manual(values = c(`ewing_sarcoma`=0.05,`neuroblastoma`= 0.025,`embryonal_rhabdomyosarcoma` = 0.05, `nephroblastoma` = 0.03, `hepatoblastoma` = 0.1, `alveolar_rhabdomyosarcoma` = 0.18),guide = "none") +
  scale_alpha_manual(values=c(`unknown`=0.8,`driver`=1,`target_gene_cn_change`=1,chrom_alt=1),guide = "none")

p = p + facet_grid(cols=vars(cancer_type),scales="free",space = "free",labeller = labeller(cancer_type = cancer_type.labs))

p + theme(legend.position="bottom")#, legend.direction = "vertical")


p +
  geom_point(data=filter(display_patient_level_complex_sv_overview,driver_gene_mutated),aes(x=patient_label,y=-0.25,shape="clinrel")) +
  geom_point(data=filter(display_patient_level_complex_sv_overview,tp53_or_mdm2),aes(x=patient_label,y=-0.5,shape="tp53/mdm2")) +
  geom_point(data=filter(display_patient_level_complex_sv_overview,clinical_event),aes(x=patient_label,y=-0.75),color="red3") 

p + theme(legend.position="bottom")  +
  geom_point(data=filter(display_patient_level_complex_sv_overview,driver_gene_mutated),aes(x=patient_label,y=-0.25,shape="clinrel")) +
  geom_point(data=filter(display_patient_level_complex_sv_overview,tp53_or_mdm2),aes(x=patient_label,y=-0.5,shape="tp53/mdm2")) +
  geom_point(data=filter(display_patient_level_complex_sv_overview,clinical_event),aes(x=patient_label,y=-0.75),color="red3") 


p +
  geom_point(data=filter(display_patient_level_complex_sv_overview,driver_gene_mutated_no_complex),aes(x=patient_label,y=-0.25,shape="clinrel_no_complex")) +
  geom_point(data=filter(display_patient_level_complex_sv_overview,tp53_or_mdm2),aes(x=patient_label,y=-0.5,shape="tp53/mdm2")) +
  geom_point(data=filter(display_patient_level_complex_sv_overview,clinical_event),aes(x=patient_label,y=-0.75),color="red3") 



p = ggplot(display_complex_svs_annotated,aes(x=patient_label,pattern=effect)) + 
  geom_bar_pattern(pattern_color="black",color="white",pattern_fill="black") +
  scale_pattern_manual(values=annotation_pattern_effect)
p
p + theme(legend.position="bottom")#, legend.direction = "vertical")

dev.off()
```

# Manuscript numbers

```{r}
clinrel_mutations_incl_prognostic  %>% uq_patients()
clinrel_mutations_incl_prognostic %>% filter(sv_alteration)  %>% uq_patients()
clinrel_mutations_incl_prognostic %>% filter(complex_sv_alteration) %>% uq_patients()

patient_level_complex_sv_overview %>% filter(cancer_type=="ewing_sarcoma") %>% uq_patients()
patient_level_complex_sv_overview %>% filter(complex_sv_driver_gene_expected& cancer_type=="ewing_sarcoma") %>% uq_patients()

patient_level_complex_sv_overview %>% filter(complex_svs_cnt_driver>0 & complex_svs_cnt_driver==complex_svs_cnt) %>% uq_patients()
patient_level_complex_sv_overview %>% filter(complex_sv_driver_gene_expected
                                     & complex_svs_cnt_driver>0 & complex_svs_cnt_effect==complex_svs_cnt
                                     ) %>% uq_patients() # M157AAB = 19 of 21
#for 19 of 21 tumors with expected ddriver complex sv,it is only complex sv

patient_level_complex_sv_overview %>% filter(!complex_sv_driver_gene_expected
                                     & complex_svs_cnt_effect>0 & complex_svs_cnt_effect==complex_svs_cnt
                                     ) %>% uq_patients() #for 10 of 21 tumors with candidate complex sv,it is only complex sv


patient_level_complex_sv_overview %>% filter(complex_svs_cnt_driver>0) %>% uq_patients()
patient_level_complex_sv_overview %>% filter(!complex_sv_driver_gene_expected & complex_svs_cnt_driver>0) %>% uq_patients()
patient_level_complex_sv_overview %>% filter(complex_svs_cnt_driver==0 & complex_svs_cnt_effect>0) %>% uq_patients()
patient_level_complex_sv_overview %>% filter(complex_svs_cnt_effect>complex_svs_cnt_driver & complex_svs_cnt_driver>0) %>% uq_patients()

complex_svs_annotated %>% 
  group_by(effect) %>%
  summarise(complex_cnt=n(),patient_cnt=cnt_str(patient_label))

```
breakpoints nearby cancer genes
```{r}

svs_per_superclass = merged_svs %>% cnt_uq(attr_name="svs_cnt",group_cols = c("complex_sv_class_super"))

#only consider coding genes otherwise very high
svs_per_superclass_summary = svs_per_superclass %>% 
  left_join(merged_svs %>% filter(flag_cancer_gene_sv_bp_1mb_pmc)  %>% cnt_uq(attr_name="svs_cnt_gene_1mb_pmc",group_cols = c("complex_sv_class_super")) ) %>%
  left_join(merged_svs  %>% filter(flag_cancer_gene_sv_bp_pmc) %>% cnt_uq(attr_name="svs_cnt_in_gene_pmc",group_cols = c("complex_sv_class_super"))) %>%
  left_join(merged_svs %>% filter(flag_cancer_gene_sv_bp_1mb)  %>% cnt_uq(attr_name="svs_cnt_gene_1mb_cancer_gene",group_cols = c("complex_sv_class_super")) ) %>%
  left_join(merged_svs  %>% filter(flag_cancer_gene_sv_bp) %>% cnt_uq(attr_name="svs_cnt_in_gene_cancer_gene",group_cols = c("complex_sv_class_super")))

total_sums = svs_per_superclass_summary[2:6] %>% colSums() %>% t() %>% as.data.frame("total") 
total_sums$complex_sv_class_super=rownames(total_sums)

complex_sums = svs_per_superclass_summary[1:4,2:6] %>% colSums() %>% t() %>% as.data.frame("all_complex") 
complex_sums$complex_sv_class_super=rownames(complex_sums)

svs_per_superclass_summary = rbind(svs_per_superclass_summary,complex_sums,total_sums)
svs_per_superclass_summary =  svs_per_superclass_summary %>% 
  dplyr::mutate(
                frac_gene_1mb_pmc = svs_cnt_gene_1mb_pmc/svs_cnt,
                frac_in_gene_pmc = svs_cnt_in_gene_pmc/svs_cnt,
                frac_gene_1mb_cancer = svs_cnt_gene_1mb_cancer_gene/svs_cnt,
                frac_in_gene_cancer = svs_cnt_in_gene_cancer_gene/svs_cnt)

svs_per_superclass_summary

merged_svs$patient_sv_merged %>% unique() %>% length()
merged_svs %>% filter(flag_is_complex) %>% select(patient_sv_merged) %>% unique() %>% nrow()
svs_per_superclass_summary %>% select(complex_sv_class_super,svs_cnt,contains("pmc"))
```


```{r}
complex_sv_superclass_patients = merged_svs %>% cnt_uq(attr_name="patient_cnt",group_cols = c("complex_sv_class_super"),uq_id = "patient_label")

complex_sv_superclass_patients
```
